GPU-based Adaptive Beamforming for Active Sonar Imaging

### 2012-05-08 ECUA2012/gpu_capon_ecua2012.tex
ECUA contribution almost ready.

\begin{figure}[!t]
\centering
\includegraphics[width=\linewidth]{gfx/benchmark.pdf}
\caption{\gls{MVDR} benchmarks. Forming a 55\,kpixels image from a $M=32$ channel array.}\label{benchmarks}
\end{figure}

The other dataset was a 55\,kpixels sectorscan test image. We used it to benchmark our \gls{MVDR} implementation, and compared the result with the runtimes of an optimized single-thread C and Matlab implementation (\Fig{benchmarks}). Processing times are shown for varying subarray lengths $L$ and varying amounts of temporal averaging $N_\text{avg}$. The test system was a quad-core Intel Xeon E5620, with 64Gb of \gls{RAM}, and an nVidia Quadro 6000 card. This is a \gls{GPU} with 448 \gls{CUDA} cores and 6Gb of onboard \gls{RAM}, capable of performing roughly 1Tflops.

@@

The \gls{MVDR} beamformer is an algorithm well suited for implementation on a \gls{GPU}. This is because the computations involved are independent on a pixel level, and partially also within each pixel. We were able to achieve speedup factors of 2-3 orders of magnitude by implementing our active sonar capable \gls{MVDR} implementation on a high-end \gls{GPU}. This performance is sufficient for computing critically sampled full-coverage sectorsscan images from the HISAS1030 sonar in realtime. Furthermore, even if such processing speeds are not required, it might be advantageous to relieve the \gls{CPU} from some of its workload. After all, these chips are designed to cooperate.

>>>>

\begin{figure}[!t]
\centering
\includegraphics[width=\linewidth]{gfx/benchmark.pdf}
\caption{\gls{MVDR} benchmarks. Forming a 55\,kpixel image from a $M=32$ channel array.}\label{benchmarks}
\end{figure}

The other dataset was a 55\,kpixel sectorscan test image. We used it to benchmark our \gls{MVDR} implementation, and compared the result with the runtimes of an optimized single-thread C and Matlab implementation (\Fig{benchmarks}). Processing times are shown for varying subarray lengths $L$ and varying amounts of temporal averaging $N_\text{avg}$. The test system was a quad-core Intel Xeon E5620, with 64Gb of \gls{RAM}, and an nVidia Quadro 6000 card. This is a \gls{GPU} with 448 \gls{CUDA} cores and 6Gb of onboard \gls{RAM}, capable of performing roughly 1Tflops.

@@

The \gls{MVDR} beamformer is an algorithm well suited for implementation on a \gls{GPU}. This is because the computations involved are independent on a pixel level, and partially also within each pixel. We were able to achieve speedup factors of 2-3 orders of magnitude by implementing our active sonar capable \gls{MVDR} implementation on a high-end \gls{GPU}. This performance is sufficient for computing critically sampled full-coverage sectorsscan images from the HISAS1030 sonar in realtime. Furthermore, even if such processing speeds are not required, it might be advantageous to relieve the \gls{CPU} from some of its workload. After all, these chips were designed to cooperate.
